Learning Graphical Concepts

### 2013-10-23 drafts/nips/nips2013.tex
writeup

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{clrscode4e}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{natbib}

@@

\title{Incremental Construction of Programs in a Multitask Setting}

@@

One example of a structured representation we would like to be able to learn is the probabilistic graphical model. Although graphical models are widely used in machine learning to represent the factorized structure of probability distributions, they are generally designed by researchers rather than by algorithms. This is especially true when the goal is to learn about latent variables underlying the observed data, because it is difficult to capture the knowledge which allows researchers to make reasonable assumptions about which latent variables exists and what they depend on.

In particular, machine learning researchers make frequent use of graphical motifs and abstractions such as trees, chains, rings, grids, mixtures, and plates to constrain the space of graphical models they consider. But how do they learn that these are good motifs? How do they learn how to use them and combine them to design new models?

One hypothesis is that these motifs and abstractions are best represented as programs that manipulate and compose graph operations. In this work, we discuss how one could learn these motifs and abstractions by finding programs that generate common graphical model types. In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm, that attempts to solve many related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs. We discuss how SEC can be used to extract from a set of training tasks the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.

@@

The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks. Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
%This distribution helps guide its search to relevant parts of of the space of programs.
In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs that are likely to be relevant for the tasks it has been given; and second, compresses the solutions it found and uses this compression generate a new weighted library of primitives that defines an updated $\mathcal{D}$.

@@

@@

\begin{figure}
\begin{codebox}
\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Draw grammar from description-length prior}
\li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
\li \Comment{Draw $\rho_n$, the $n^{th}$ program, by composing its subprograms, $e_n^{i}$}
\li \id{e_n^{i}} $\sim P_{e|G}(\cdot | G)$
\li \id{\rho_n} $= e_n^{\ell_n} (e_n^{\ell_n-1} ( \cdots e_n^{1}))$
\li \Comment{Draw $t_n$ by adding noise $\epsilon_n$ to the output of $\rho_n$}
\li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
\li \id{t_n} $= \rho_n + \epsilon_n$
\end{codebox}
\caption{The generative model underlying the SEC algorithm. \label{ref:gm}}
\end{figure}

@@

Exact computation of the normalizing constant in equation (\ref{eq:qdist}), or the expectation in equation (\ref{eqn:gmax}), would require summing over the infinite space of programs that can be generated from the grammar. To approximate this expectation, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
%Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $e'$, and apply that new function to produce the program $e'(e)$.
Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.

In our experiments, we used beam search. The objective function was the unnormalized posterior $P_{t_n|\rho_n}(t_n | \cdot )P_{\rho_n | G}(\cdot | G)$.
The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
Similarly, the search moves considered were those functions with the most highest prior probability under $G$.

Equation (\ref{eq:gmax}) has a natural interpretation as a form of compression.
Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.

Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
We approximate the description length of the grammar by the number of unique program fragments in the grammar, multiplied by a regularization constant.
The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
A slight generalization of the Neville-Manning algorithm then permits tractible approximate minimization of (\ref{eq:gmax}).
Estimation of the production probabilities of $G$ is performed as in \cite{DBLP:conf/ijcai/DechterMAT13}.

\subsection{Program Representation}

Following~\citet{DBLP:conf/ijcai/DechterMAT13} and~\citet{DBLP:conf/icml/LiangJK10}, we use a
simply typed combinatory calculus -- a variable-free subset of the
polymorphic simply-typed lambda calculus~\citep{DBLP:books/daglib/0005958} -- as our
program representation language. The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple. A more detailed explanation and discussion of this program representation can be found in the papers cited above.

@@

  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
  \caption{Example input/output pairs of learned programs. Each induced program takes as input an arbitrarily large list of observed nodes and produces an undirected graphical model. Labeled/Unlabeled nodes correspond to observed/latent variables.}
  \label{fig:tasks}
\end{minipage}
\hspace{0.2in}
\begin{minipage}[t]{.5\textwidth}
  \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
  \caption{Union: computes union of vertexes and edges. UnionC: adds edges between corresponding vertexes. UnionHT: adds edge between first (head) vertex and last (tail) vertex. UnionD: relabels input graph vertices so that they do not overlap and then computes union.}
  \label{fig:graphcomb}\par\vfill
% \end{subfigure}
\end{minipage}
\end{figure}

\subsection{Experiment}

@@

The SEC algorithm was provided with an initial library of primitive  combinators, which included the basis combinators of the combinatory calculus, function list operations such as \textsc{Map} and \textsc{Foldr}/\textsc{Foldl}. We also included four binary operations on undirected graphical models, each of which computes some kind of union between two graphs. These graph operations are diagrammed in Figure~\ref{fig:graphcomb}.

@@

We ran the SEC algorithm on the set of twelve tasks. The algorithm is parameterized by the frontier size, the number of unique programs drawn from the library on each iteration. We ran the algorithm both for frontier sizes of 1000 and 5000. Note that because proposed solutions are compositions of these programs, the algorithm looks at many more potential solutions than the number of programs in the frontier.

With a frontier size of 1000, the algorithm solves 7/12 tasks. With a frontier size of 5000, the algorithm solves 8/12 tasks. In each case, the algorithm returns solutions that partially satisfy the remaining tasks. As an example, we show in Figure~\ref{ref:cylinder} SEC's solution to the \textsc{Cylinder} task.

Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating $\approx 4$ million programs through its heuristic search.
However, it is able to find programs that generate simpler graphical models, such as Markov chains.
By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more successful programs.

\begin{figure}
% \begin{minipage}[b]{.45\textwidth}
%\begin{subfigure}[b]{.45\textwidth}
\includegraphics[width=0.45\linewidth]{./figures/cylinder.pdf}
\caption{The program generates a cylinder from nodes by generating two rings -- one latent, one visible -- and connecting them.}
% \end{minipage}
%\end{subfigure}
\begin{comment}
\hspace{0.2in}
\begin{minipage}[b]{.5\textwidth}
\begin{subfigure}[t]{\textwidth}
\includegraphics[width=.5\linewidth]{./figures/cylinderCombinator.pdf}
\caption{In the combinatory calculus. }
\vspace{.3in}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\begin{lstlisting}[language=haskell, basicstyle=\small, escapechar=+]
+$\lambda$+vs. let vs' = foldr unionC null vs
          vs'' = unionHT vs' (I vs')   
          vs''' = flip vs''
       in unionC vs'' vs'''
\end{lstlisting}
\vfill
\caption{The combinatory calculus expression tranlated into \textsc{Haskell}-like syntax for readability.}
% \end{minipage}
\end{subfigure}
\end{minipage}
\end{comment}
%\caption{An example of solution SEC finds to solve \textsc{Cylinder}\label{ref:cylinder}}
\end{figure}

What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
%Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
After one iteration, the program fragment \verb|(foldr unionC)|, used in the Markov chain program, is incorporated in to the grammar.
On the second iteration, the Ising model is learned using this fragment, producing the program \verb|(((S B) map) ((foldr unionC) null))|.

@@

We have presented the SEC algorithm in the context of learning graph structures, but we note that SEC is a general purpose multitask program induction algorithm. For an excellent survey of the current state of inductive programming systems, see~\citep{DBLP:conf/aaip/Kitzelmann09}.

>>>>

\documentclass{article} % For LaTeX2e
\usepackage{nips13submit_e,times}
\usepackage{hyperref}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{clrscode4e}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage{wrapfig}

@@

\title{Learning Graphical Concepts}

@@

One example of a structured representation we would like to be able to learn is the probabilistic graphical model. %Although graphical models are widely used in machine learning
% to represent the factorized structure of probability distributions
% , they are generally designed by researchers rather than by algorithms.
%This is especially true when the goal is to learn about latent variables underlying the observed data, because it is difficult to capture the knowledge which allows researchers to make reasonable assumptions about which latent variables exists and what they depend on. 
Machine learning researchers make frequent use of graphical motifs and abstractions such as trees, chains, rings, grids, mixtures, and plates to constrain the space of graphical models they consider. But how do they learn that these are good motifs? How do they learn how to use them and combine them to design new models?

One hypothesis is that these motifs and abstractions are best represented as programs that manipulate and compose graph operations. In this work, we discuss how one could learn these motifs and abstractions by finding programs that generate common graphical models. In particular, we present the SEC algorithm, a general purpose multitask program induction algorithm.
%, that attempts to solve related tasks by generating a weighted library of useful subroutines that induces a distribution over the space of programs.
We discuss how SEC can be used to extract from a set of examples the kind of abstract knowledge that machine learning researchers bring to the task of designing graphical models, and we speculate that learning such representations can assist in the difficult tasks of automatically finding good latent-variable graphical models for novel datasets.

@@

The SEC algorithm takes a library of primitives and a set of related tasks to solve; its goal is to produce a set of programs that solve these tasks. Like the EC algorithm~\cite{DBLP:conf/ijcai/DechterMAT13}, of which it an extension, the SEC algorithm maintains a distribution $\mathcal{D}$ over programs.
%This distribution helps guide its search to relevant parts of of the space of programs.
In each iteration, SEC does two things: first, it uses $\mathcal{D}$ to explore the space of programs
%that are likely to be relevant for the tasks it has been given
; and second, compresses the solutions it found to generate a new weighted library of primitives that define an updated $\mathcal{D}$.

Following~\citet{DBLP:conf/ijcai/DechterMAT13} and~\citet{DBLP:conf/icml/LiangJK10}, SEC learns programs expressed in
simply typed combinatory calculus 
%-- a variable-free subset of the
%polymorphic simply-typed lambda calculus
~\citep{DBLP:books/daglib/0005958} %-- as our program representation language.
 The simplicity the combinatory calculus makes describing a distribution over its expressions particularly simple. A more detailed explanation and discussion of this program representation can be found in the papers cited above.

@@

\begin{wrapfigure}{R}{0.6\textwidth}
\begin{codebox}
\li \id{G} $\sim$ $P_G(\cdot)$ \Comment{Description-length prior}
\li \id{\ell_n} $\sim\mbox{Uniform}(1, L)$ \Comment{Draw number of subprograms in $n^{th}$ program}
\li \Comment{Draw $\rho_n$, the $n^{th}$ program, by composing its subprograms, $e_n^{i}$}
\li \id{e_n^{i}} $\sim P_{e|G}(\cdot | G)$
\li \id{\rho_n} $= e_n^{\ell_n} (e_n^{\ell_n-1} ( \cdots e_n^{1}))$
\li \Comment{Draw $t_n$ by adding noise $\epsilon_n$ to the output of $\rho_n$}
\li \id{\epsilon_n} $\sim P_{\mbox{noise}}(\cdot)$
\li \id{t_n} $= \textsc{Eval}(\rho_n) + \epsilon_n$
\end{codebox}
\caption{The generative model underlying the SEC algorithm. \label{ref:gm}}
\end{wrapfigure}

@@

@@

Exact computation of the normalizing constant in equation (\ref{eq:qdist}), or the expectation in equation (\ref{eqn:gmax}), would require summing over the infinite space of programs that can be generated from the grammar. To approximate this expectation, we perform a heuristic search over the space of programs, with the goal of finding those programs for which $q(\cdot)$ is high.
%Rather than use mutation and crossover operators, as done in genetic programming, or random resampling of subprograms, as done in recent work in applying Metropolis-Hastings to program learning, we propose that the search moves should themselves be programs drawn from the grammar $G$.
%Concretely, in order to modify a program $e$ to produce new candidate programs, we draw a new function from our grammar, $e'$, and apply that new function to produce the program $e'(e)$.
%Thus, we use the grammar not only as a distribution over programs of interest, but also as a distribution over programs that take as input other programs.

In our experiments, we used beam search. The objective function was equation (\ref{eq:qdist}).
The beam was initialized to be the programs with the highest prior probability according to the grammar $G$.
Similarly, the search moves considered were those functions with the most highest prior probability under $G$.

Equation (\ref{eq:gmax}) has a natural interpretation as a form of compression.
Interpreting negative log probabilities as description lengths, the update (\ref{eq:gmax}) picks the grammar minimizing the sum of the description length of the grammar, plus the expected description length of the programs found to solve the tasks.
A slight generalization of the Neville-Manning algorithm permits tractible approximate minimization of (\ref{eq:gmax}).

\begin{comment}
Performing the minimization operation in equation (\ref{eq:gmax}) is generally difficult.
We approximate the description length of the grammar by the number of unique program fragments in the grammar, multiplied by a regularization constant.
The description length of a program is approximated by the number of (not necessarily unique) program fragments in its smallest parse tree given the grammar.
A slight generalization of the Neville-Manning algorithm then permits tractible approximate minimization of (\ref{eq:gmax}).
Estimation of the production probabilities of $G$ is performed as in \cite{DBLP:conf/ijcai/DechterMAT13}.
\end{comment}

@@

  \includegraphics[width=\linewidth]{./figures/tasks.pdf}
  \caption{Example input/output pairs of learned programs. Each induced program takes as input an arbitrarily large list of observed nodes and produces an undirected graphical model. Labeled/Unlabeled nodes correspond to observed/latent variables.}
  \label{fig:tasks}
\end{minipage}
\hspace{0.2in}
\begin{minipage}[t]{.5\textwidth}
  \includegraphics[width=\linewidth]{./figures/GraphCombinators.pdf}
  \caption{Union: computes union of vertexes and edges. UnionC: adds edges between corresponding vertexes. UnionHT: adds edge between first (head) vertex and last (tail) vertex. UnionD: relabels input graph vertices so that they do not overlap and then computes union.}
  \label{fig:graphcomb}\par\vfill
% \end{subfigure}
\end{minipage}
\vspace{-0.5cm}
\end{figure}

\subsection{Experimental Setup}

@@

The SEC algorithm was provided with an initial library of primitive  combinators, which included the basis combinators of the combinatory calculus, function list operations such as \textsc{Map} and \textsc{Foldr}/\textsc{Foldl}. We also included four binary operations on undirected graphical models, each of which computes a union between two graphs. These graph operations are diagrammed in Figure~\ref{fig:graphcomb}.

@@

We ran the SEC algorithm on the set of twelve tasks. The algorithm is parameterized by the frontier size, the number of unique programs drawn from the library on each iteration. %We ran the algorithm both for frontier sizes of 1000 and 5000. %Note that because proposed solutions are compositions of these programs, the algorithm looks at many more potential solutions than the number of programs in the frontier. 
With a frontier size of 1000, the algorithm solves 7/12 tasks. With a frontier size of 5000, the algorithm solves 8/12 tasks. In each case, the algorithm returns solutions that partially satisfy the remaining tasks. Figure~\ref{ref:cylinder} shows SEC's solution to the \textsc{Cylinder} task. 
\begin{wrapfigure}{R}{0.6\textwidth}
\includegraphics[width=\linewidth]{./figures/cylinder.pdf}
\caption{The program generates a cylinder from nodes by generating two rings -- one latent, one visible -- and connecting them.\label{ref:cylinder}}
\end{wrapfigure}
Initially, SEC is unable to find solutions to the complicated graph structure tasks, such as the Ising Model or the Hidden Markov Model, even after enumerating $\approx 4$ million programs through its heuristic search.
However, it is able to find programs that generate simpler graphical models, such as Markov chains.
By compressing the solutions to these simpler problems, it modifies its search procedure and, in subsequent iterations, learns more successful programs.

What grammars does SEC learn, and how do they enable the bootstrap learning of more complicated programs?
The simplest example is that of how the Markov chain bootstraps the learning of the Ising model.
%Most of the easy tasks, such as the Markov chain, can be solved with the right choice of one higher-order function and one undirected union.
After one iteration, the program fragment \textsc{(foldr unionC)}, used in the Markov chain program, is incorporated in to the grammar.
On the second iteration, the Ising model is learned using this fragment, producing the program
\textsc{(((S B) map) ((foldr unionC) null))}.

@@

%We have presented the SEC algorithm in the context of learning graph structures, but we note that SEC is a general purpose multitask program induction algorithm. For an excellent survey of the current state of inductive programming systems, see~\citep{DBLP:conf/aaip/Kitzelmann09}.
